Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone

### 2020-08-28 paper/ifo-complexAudio.tex
editing the future work and conclusions

where $P_{xx}(f),P_{ss}(f),P_{ww}(f)$ stands for the frequency spectrum of the noisy observation, clean signal, and pure noise respectively. 

@@

The choice of $n=100$ is a reasonably selection to balance smoothing and efficiency.

The timeseries and frequency spectrum results of the Wiener fileter are are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively. 
There is improvement which is particularly visible in the timeseries, however a strong noise hum persists. 
%The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 
%It minimizes the mean square error $\mathbb{E}[(\hat{x}-x)^2]$ between $\hat{x}$ and the actual signal $x$~\citep{verhoeven2009robust}.
%\han{should we add more details here too?} \jam{[seconded]}
%The timeseries and frequency spectrum results of the Wiener filter are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively. 
%The timeseries is visibly cleaner, however there is a strong noise hum.

@@

Speech enhancement for a noisy channel is a classic problem in signal processing. 
Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compare 13 speech enhancement methods and find the best to be a log minimum mean-square error (log-MMSE) estimator. 
This type of estimator is based on work by Ephraim~\&~Malah~(1984)~\cite{Ephraim1984SpeechEU_logMMSE} into speech enhancement by minimising the mean square error (MSE) to the logarithm of the spectral (Fourier) amplitude.
The logarithm approximates the response of the human ear~\cite{SubjectiveComparison}.

In this work we use an existing implementation~\cite{logmmse} of the logMMSE estimator. 
Applying this estimator to the optical microphone output produced dramatically cleaner results.

In Figures~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum} we present results from playing a speech recording through the optical microphone. 
Figure~\ref{fig:logMMSE_timeseries} shows the anti-aliased photodiode time series from before (top panel) and after (bottom panel) applying the logMMSE filter. 
Figure~\ref{fig:logMMSE_spectrum} shows the corresponding spectra before (top panel) and after (bottom panel) the logMMSE filter. 
Comparing the before and after spectra shows significant attenuation of all mains harmonics as well as a general smoothing of the spectrum.

@@

We also find improvement with music recordings.
After applying the estimator, simple chords and drums can be heard clearly. 
However more complex melodies are sometimes completely absent. 
From observation, this is especially true for instruments of certain timbres (harmonic profiles), in particular flutes and violins.
This could be a perceptual effect or a frequency dependence somewhere in the optical microphone. 
Speculating, perhaps the speaker-mirror oscillates more at at low frequencies and so instruments like electric bass and drums are louder in the results.

Overall, we find that the intelligibility of speech remains low. 
The played-back voice sounds mumbled and indistinct. 
The estimator removes most of the background noise, but does not clarify the speech. 
To address these problems with the recordings we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all. 
This requires a better understanding of the system.
% The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio. However, upon comparing injected pure tones, the increase appears to be minimal and inconsistent, at around 5Hz maximum difference. An simple explanation is that the speaker deflections are small compared to manual pressing on the optical table.

>>>>

where $P_{xx}(f),P_{ss}(f),P_{ww}(f)$ are the frequency spectrum of the noisy observation, clean signal, and pure noise respectively. 

@@

The choice of $n=100$ is a reasonable selection to balance smoothing and efficiency.

The timeseries and frequency spectrum results of the Wiener filter are are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively. 
There is improvement which is particularly visible in the timeseries, however a strong noise hum persists. 
%The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 
%It minimizes the mean square error $\mathbb{E}[(\hat{x}-x)^2]$ between $\hat{x}$ and the actual signal $x$~\citep{verhoeven2009robust}.
%\han{should we add more details here too?} \jam{[seconded]}
%The timeseries and frequency spectrum results of the Wiener filter are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively. 
%The timeseries is visibly cleaner, however there is a strong noise hum.

@@

Speech enhancement for a noisy channel is a classic problem in signal processing. 
In Ref.~\cite{SubjectiveComparison}, $13$ speech enhancement methods are compared and the best found to be a log minimum means-square error (logMMSE) estimator. 
This type of estimator is based on work in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} into speech enhancement by minimising the mean square error (MSE) to the logarithm of the spectral (Fourier) amplitude.
%Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compare 13 speech enhancement methods and find the best to be a log minimum mean-square error (log-MMSE) estimator. 
%This type of estimator is based on work Ephraim~\&~Malah~(1984)~\cite{Ephraim1984SpeechEU_logMMSE} into speech enhancement by minimising the mean square error (MSE) to the logarithm of the spectral (Fourier) amplitude.
The logarithm approximates the response of the human ear~\cite{SubjectiveComparison}.
We apply an existing implementation of the logMMSE estimator (see Ref.~\cite{logmmse}).

%In this work we use an existing implementation~\cite{logmmse} of the logMMSE estimator. 
%Applying this estimator to the optical microphone output produced dramatically cleaner results.

The timeseries and frequency results of the logMMSE are shown in Figs.~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum} respectively. 
The top two panels of each figure again show the original speech signal and anti-aliased reading from the optical microphone for ease of comparison. 
The third panel of Figs.~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum} show the timeseries and corresponding spectra after applying the logMMSE filter.

@@

The results show significant attenuation of the mains harmonics as well as a general smoothing of the spectrum.
However, we find that the intelligibility of speech remains low with the voice sounding mumbled and indistinct. 
The logMMSE removes most of the background noise, but does not clarify the speech.

We find some improvement with music recordings. 
After applying the logMMSE, simple chords and drums can be heard clearly, however more complex melodies are sometimes completely absent. 
From observation, this is especially true for instruments of certain timbres (harmonic profiles), in particular flutes and violins.
This could be a perceptual effect or a frequency dependence somewhere in the optical microphone. 
Speculating, perhaps the speaker-mirror oscillates more at at low frequencies and so instruments like electric bass and drums are louder in the results.

To address these problems we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all. 
This requires a better understanding of the system as discussed in Section~\ref{sec:future_work}.

%Overall, we find that the intelligibility of speech remains low. 
%The played-back voice sounds mumbled and indistinct. 
%The estimator removes most of the background noise, but does not clarify the speech. 
%To address these problems with the recordings we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all. 
%This requires a better understanding of the system.
% The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio. However, upon comparing injected pure tones, the increase appears to be minimal and inconsistent, at around 5Hz maximum difference. An simple explanation is that the speaker deflections are small compared to manual pressing on the optical table.

### 2020-08-28 paper/ifo-conclusions.tex
editing the future work and conclusions

In this paper we present an educational demonstration in which a table top Michelson interferometer serves as an analogue to a gravitational-wave detector. 
Vibrations produced by gravitational-wave signals are simulated with sound recordings played though a speaker fixed to the back of one of the interferometer mirrors. 
We consider quasi-monochromatic signals as an analogue to continuous wave signals as well as short segments of speech and music which are (more distant) analogues of unmodelled gravitational-wave bursts, e.g. from supernovae. 
% This serves to demonstrate signal processing techniques.

@@

In Section~\ref{sec:optical_microphone} we demonstrate recovery of more complex signals of music and speech. 
Substituting the webcam for the increased frequency response of a photodiode read by a Raspberry Pi, we transform the interferometer into an optical microphone.
We demonstrate that the optical microphone is capable of playing back simple recordings well but that it produces unintelligible recordings of speech. 
Mains hum at $50\,{\rm Hz}$ and its harmonics dominates the photodiode output. 
This offers students the opportunity to explore speech enhancement techniques to remove it, including the well performing logMMSE estimator. 
Further avenues of exploration include a more complete model of the optical microphone transfer function, which may lead to improvements in intelligibility.

The demonstration presented here has good potential to be adapted for use in the undergraduate laboratory. 
It can be used to demonstrate topics of interest to physicists and electrical engineers including: interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement. 
The detection of gravitational waves in 2015 led to increased excitement and public interest in gravitational wave science. 
The table-top demonstration presented here can also be adapted as a tool to explain gravitational wave research developments for a wider non-specialist audience. 
The future will bring many more detections, including further observations of transient signals from binary black hole and binary neutron star mergers, as well as the anticipated first detections of other classes of signal, including continuous waves.

>>>>

In this paper we present an educational demonstration in which a table-top Michelson interferometer serves as an analogue to a gravitational-wave detector. 
Vibrations produced by gravitational-wave signals are simulated with sound recordings played though a speaker fixed to the back of one of the interferometer mirrors. 
We consider quasi-monochromatic signals as an analogue to continuous wave signals as well as short segments of speech and music which are (more distant) analogues of unmodelled gravitational-wave bursts, e.g. from supernovae. 
% This serves to demonstrate signal processing techniques.

@@

In Section~\ref{sec:optical_microphone} we demonstrate recovery of more complex signals of music and speech. 
Substituting the webcam for the increased frequency response of a photodiode read by a Raspberry Pi, we transform the interferometer into an optical microphone.
We demonstrate that the optical microphone is capable of playing back simple recordings well but produces unintelligible recordings of speech. 
Mains hum at $50\,{\rm Hz}$ and its harmonics dominates the photodiode output. 
This demonstration offers students the opportunity to explore a variety of filtering and speech enhancement techniques. 
Further avenues of exploration include a more complete model of the optical microphone transfer function, which may lead to improvements in intelligibility.

The demonstration presented here has good potential to be adapted for use in the undergraduate laboratory. 
It can be used to demonstrate topics of interest to physicists and electrical engineers including: interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement. 
Given the increased excitement and public interest in gravitational-wave science since the detection in 2015, this table-top demonstration presented here can also be adapted as a tool to explain gravitational-wave research developments for a wider non-specialist audience. 
The future will bring many more detections, including further observations of transient signals from binary black hole and binary neutron star mergers, as well as the anticipated first detections of other classes of signal, including continuous waves.

### 2020-08-28 paper/ifo-futureWork.tex
editing the future work and conclusions

An initial improvement is to better isolate the experiment from interference, such as the mains supply. 
The Raspberry Pi is USB powered and requires low power so it could be run off a commercial external phone battery. 
Everything else in the circuit, including the photodiode, is connected to the Pi. 
The laser is DC powered through a converter, so should not introduce mains harmonics itself. 
Transferring the circuit from a breadboard to a printed circuit board will also reduce interference~\cite{elfekey2013design}

@@

A second improvement is to measure the full model of the transfer function of the system, providing more information about how the audio signal couples with the interferometer.
% commented out whilst James is checking this
%With perfect knowledge of the transfer function, signal recovery is limited by the signal-to-noise ratio as an optimal matched filter could be constructed.
%\jam{[check with Rob et al. about the veracity of this]}
The transfer function starts from the voltage sent to the speaker and ends with the voltage recorded by the Pi. It includes any non-linearities in the speaker itself, the speaker-mirror coupling, the path length to intensity relation (Appendix~\ref{app:intensity_derivation}), and the photodiode reading. 
A schematic of the full optical microphone transfer function is shown in Figure~\ref{fig:pipeline_highlighted}.

A voluminous literature exists about using a hidden Markov model (HMM) trained against phonemes to recognize speech~\cite{HMM_english}. 
Interestingly, this connects back to the Viterbi algorithm in Section IV, which is also underpinned by a hidden Markov model, albeit of a different type. 
Alternatively, machine learning solutions exist throughout the field that are able to compete~\cite{SEGAN} with statistical techniques like the logMMSE estimator.

Other traditional signal processing methods for extracting spectrogram like wavelet transform (WT) \citep{nason1995stationary} can also be implemented and compared with HMM. WT provides both time and frequency information, making it easier to pinpoint the origin of noise respect to time. 
In Refs.~\cite{tufekci2000feature,agbinya1996discrete}, WT was also proposed for speech recognition.

>>>>

An initial improvement is to better isolate the experiment from interference, such as the mains supply. 
The Raspberry Pi is USB powered and requires low power so it can instead be powered by a commercial external phone battery rather than the mains. 
Everything else in the circuit, including the photodiode, is connected to the Pi. 
The laser is DC powered through a converter, so should not introduce mains harmonics itself. 
Transferring the circuit from a breadboard to a printed circuit board can also reduce interference~\cite{elfekey2013design}

@@

A second improvement is to measure the full model of the transfer function of the system, providing information about how the audio signal couples with the interferometer.
% commented out whilst James is checking this
%With perfect knowledge of the transfer function, signal recovery is limited by the signal-to-noise ratio as an optimal matched filter could be constructed.
%\jam{[check with Rob et al. about the veracity of this]}
The transfer function starts from the voltage sent to the speaker and ends with the voltage recorded by the Pi. 
It includes any non-linearities in the speaker itself, the speaker-mirror coupling, the path length to intensity relation (Appendix~\ref{app:intensity_derivation}), and the photodiode reading. 
A schematic of the full optical microphone transfer function is shown in Fig.~\ref{fig:pipeline_highlighted}.

A voluminous literature exists about using a hidden Markov model (HMM) trained against phonemes to recognize speech~\cite{HMM_english}. 
Interestingly, this connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a hidden Markov model, albeit of a different type. 
Alternatively, machine learning solutions exist throughout the field that are able to compete~\cite{SEGAN} with statistical techniques like the logMMSE estimator.

Other traditional signal processing methods for extracting spectrogram-like wavelet transform (WT) \citep{nason1995stationary} can also be implemented and compared with HMM. 
WT provides both time and frequency information, making it easier to pinpoint the origin of noise respect to time. 
In Refs.~\cite{tufekci2000feature,agbinya1996discrete}, WT is also proposed for speech recognition.

### 2020-08-28 paper/paper_main.tex
editing the future work and conclusions

\begin{acknowledgments}
The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance throughout the project. 
This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004). 
This work also received financial support from the Institute of Physics International Member Grant and travel support from the Australian National University PhB Science program.

>>>>

\begin{acknowledgments}
The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance throughout the project. 
This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004). 
This work also received financial support for hardware from the Institute of Physics International Member Grant and for travel from the Australian National University PhB Science program.
